New insights into the Weddell Sea ecosystem applying a network approach

### 2022-11-14 Manuscript/Manuscript.Rmd
Finished Discussion & Conclusion

Introduction text goes here.

The objective of this work was twofold: 1) estimate the strength for each interaction in the Weddell Sea food web, and 2) determine key trophic species considering weighted and unweighted properties and the influence on the stability of the network.

@@

Our results regarding the distribution of the interaction strengths are consistent with prior theoretical and empirical studies [e.g. @McCann1998; @Neutel2002; @Emmerson2004; @Wootton2005]. Within a system of a given complexity, commonly only a few strong species interactions are present with most interactions being weak. Historically, after @May1972 installed the debate complexity-stability, the results about the interaction strength distribution have been interpreted as an explanation for the persistence of complex food webs in nature [@Bascompte2005; @Allesina2015; @Nilsson2016]. Here we showed that this pattern is also prevalent in one of the most complex empirical marine food webs that was studied so far, with 490 species and 16041 predator-prey interactions.

Strong relationship between interaction strength and trophic level. Less stronger but evident relationship with degree.

Our results show that only few species are key in terms of stability in the Weddell Sea food web. This is in concordance with studies of complex empirical food webs in marine ecosystems in the Arctic and other locations in Antarctica [@Kortsch2015; @Marina2018; @Rodriguez2022]. In the present work we identify that such key species present particular weighted and unweighted properties: 1) high mean interaction strength; 2) mid to high trophic level; 3) relatively high number of interactions; and 4) mid to low trophic similarity. The first attempt to gain insight into the role of species in the Weddell Sea applying a complex approach was that of @Jacob2011. These authors used information on body size, predatory and feeding type, and environment (habitat) to study species' roles. After performing sequential extinction simulations, @Jacob2011 found that larger bodied species can be lost without causing a direct collapse of the network topology. A major caveat of this finding, already recognized by the authors, is that population dynamics were ignored and therefore no top-down extinctions, or other indirect effects, could occur. In our study we did consider such effects by including information on the species interaction strength, which is of paramount importance when analysing the response of perturbations in ecological communities [@McCann1998; @Montoya2009; @Novak2011]. Thus, our study suggests that species with high mean interaction strength and high trophic level need to be considered with particular attention when trying to predict the effects of perturbations on the Weddell Sea ecosystem. Furthermore, this arises as a greater issue since we have identified that these species present a mid to low trophic similarity, meaning that few other species of the food web can occupy the same trophic role. In this sense, @Murphy2016's review emphasized that polar pelagic ecosystems are particularly sensitive to change due to a low functional redundancy at key trophic levels. It's important to note that we provide a more robust analysis in this regard because we have considered species from all habitats (pelagic, benthic, land-based).

Principal results:
1) First study to
      * consider the interaction dimensionality to estimate interaction strength in a food web framework

2) Asymmetric distribution of IS at the food web level; dominance of weak interactions

3) Strong relationship between interaction strength and trophic level. Less stronger but evident relationship with degree

4) Few species are important in terms of stability: discuss species characteristics

1)
2)
Dominance of weak interactions in concordance with other studies.

Within a system of a given complexity, commonly only a few strong species interactions are present with most interactions being weak (e.g. Bascompte et al., 2005; Berlow et al., 2004; Paine, 1992; Wootton & Emmerson, 2005), though still being important for system stability (e.g. Kadoya et al., 2018; McCann et al., 1998; O'Gorman & Emmerson, 2009). Moreover, there is some evidence indicating that generalists (i.e. species with many possible interacting partners) mostly have weak interactions, while specialist species (i.e. fewer possible interacting partners) show stronger interactions (Wootton & Stouffer, 2016).

3)

4)

Our results contradict studies suggesting that mid-trophic level species are key in marine polar ecosystems. [@Murphy2016], studies by Australians

>>>>

The objective of this work was threefold: 1) estimate the strength for each interaction in the Weddell Sea food web, 2) characterise species considering weighted and unweighted properties, and 3) analyse the species' role in the stability of the food web.

@@

Our results regarding the distribution of the interaction strengths are consistent with prior theoretical and empirical studies [e.g. @McCann1998; @Neutel2002; @Emmerson2004; @Wootton2005]. Within a system of a given complexity, commonly only a few strong species interactions are present with most interactions being weak. Historically, after @May1972 installed the debate complexity-stability, the results about the distribution of the interaction strength has been interpreted as an explanation for the persistence of complex food webs in nature [@Bascompte2005; @Allesina2015; @Nilsson2016]. Here we showed that this pattern is also prevalent in one of the most complex empirical marine food webs that was studied so far, with 490 species and 16041 predator-prey interactions.

Our approach to study species properties follows @Kortsch2021 recommendation that suggests to employ a range of descriptors from both unweighted and weighted food web in order to characterise the dynamic and multifaceted nature of structural and functional ecosystem changes. We went a step forward and analysed the relationship between weighted and unweighted descriptors. In this regard, our results show a positive relationship between interaction strength (weighted) and trophic level (unweighted), and between interaction strength and degree (total number of interactions). The former relationship might contradict those studies that suggest that mid-trophic level species are involved in the major pathways of energy flow in high-latitude marine ecosystems [@Pinkerton2014; @Murphy2016; @McCormack2020; @Riccialdelli2020]. This could be explained by the lack of species biomass data in our interaction strength estimations; the methodology we applied here [@Pawar2012] allows to include empirical data for the density of the resource, though not for all food web species. Unfortunately, the lack of individual data for the entire Weddell Sea food web hampers any alternative approach. On the other hand, the positive relationship between interaction strength and degree reinforces the importance of species with many interactions: species with high degree (hubs) have a large impact on overall food web structure and functioning [@Dunne2002a; @Kortsch2015]. In the Weddell Sea, species with high degree also tend to have high mean interaction strengths. This information on the quantity and quality of interactions and its relationship enables a robust assessment of the species' role in the stability of the food web [@Cirtwill2018a].

Our results show that only few species are key in terms of stability in the Weddell Sea food web. This is in concordance with studies of complex empirical food webs in marine ecosystems in the Arctic and other locations in Antarctica [@Kortsch2015; @Marina2018; @Rodriguez2022]. In the present work we identify that such key species present particular weighted and unweighted properties: 1) high mean interaction strength; 2) mid to high trophic level; 3) relatively high number of interactions; and 4) mid to low trophic similarity. The first attempt to gain insight into the role of species in the Weddell Sea applying a complex approach was that of @Jacob2011. These authors used information on body size, predatory and feeding type, and environment (habitat) to study species' roles. After performing sequential extinction simulations, @Jacob2011 found that larger bodied species can be lost without causing a direct collapse of the network topology. A major caveat of this finding, already recognized by the authors, is that population dynamics were ignored and therefore no top-down extinctions, or other indirect effects, could occur. In our study we did consider such effects by including information on the species interaction strength, which is of paramount importance when analysing the response of perturbations in ecological communities [@McCann1998; @Montoya2009; @Novak2011]. Thus, our study suggests that species with high mean interaction strength and high trophic level need to be considered with particular attention when trying to predict the effects of perturbations on the Weddell Sea ecosystem. Furthermore, this arises as a greater issue since we have identified that these species present a mid to low trophic similarity, meaning that few other species of the food web can occupy the same trophic role. In this sense, @Murphy2016's review emphasized that polar pelagic ecosystems are particularly sensitive to change due to a low functional redundancy at key trophic levels. It's important to note that we provide a broader analysis in this regard because we have included species from all habitats (pelagic, benthic, land-based).

### 2022-11-14 Manuscript/Manuscript.tex
Finished Discussion & Conclusion

Our results regarding the distribution of the interaction strengths are
consistent with prior theoretical and empirical studies
\citep[e.g.][]{McCann1998, Neutel2002, Emmerson2004, Wootton2005}.
Within a system of a given complexity, commonly only a few strong
species interactions are present with most interactions being weak.
Historically, after \citet{May1972} installed the debate
complexity-stability, the results about the interaction strength
distribution have been interpreted as an explanation for the persistence
of complex food webs in nature
\citep{Bascompte2005, Allesina2015, Nilsson2016}. Here we showed that
this pattern is also prevalent in one of the most complex empirical
marine food webs that was studied so far, with 490 species and 16041
predator-prey interactions.

Strong relationship between interaction strength and trophic level. Less
stronger but evident relationship with degree.

Our results show that only few species are key in terms of stability in
the Weddell Sea food web. This is in concordance with studies of complex
empirical food webs in marine ecosystems in the Arctic and other
locations in Antarctica \citep{Kortsch2015, Marina2018, Rodriguez2022}.
In the present work we identify that such key species present particular
weighted and unweighted properties: 1) high mean interaction strength;
2) mid to high trophic level; 3) relatively high number of interactions;
and 4) mid to low trophic similarity. The first attempt to gain insight
into the role of species in the Weddell Sea applying a complex approach
was that of \citet{Jacob2011}. These authors used information on body
size, predatory and feeding type, and environment (habitat) to study
species' roles. After performing sequential extinction simulations,
\citet{Jacob2011} found that larger bodied species can be lost without
causing a direct collapse of the network topology. A major caveat of
this finding, already recognized by the authors, is that population
dynamics were ignored and therefore no top-down extinctions, or other
indirect effects, could occur. In our study we did consider such effects
by including information on the species interaction strength, which is
of paramount importance when analysing the response of perturbations in
ecological communities \citep{McCann1998, Montoya2009, Novak2011}. Thus,
our study suggests that species with high mean interaction strength and
high trophic level need to be considered with particular attention when
trying to predict the effects of perturbations on the Weddell Sea
ecosystem. Furthermore, this arises as a greater issue since we have
identified that these species present a mid to low trophic similarity,
meaning that few other species of the food web can occupy the same
trophic role. In this sense, \citet{Murphy2016}'s review emphasized that
polar pelagic ecosystems are particularly sensitive to change due to a
low functional redundancy at key trophic levels. It's important to note
that we provide a more robust analysis in this regard because we have
considered species from all habitats (pelagic, benthic, land-based).

Principal results: 1) First study to * consider the interaction
dimensionality to estimate interaction strength in a food web framework

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\item
  Asymmetric distribution of IS at the food web level; dominance of weak
  interactions
\item
  Strong relationship between interaction strength and trophic level.
  Less stronger but evident relationship with degree
\item
  Few species are important in terms of stability: discuss species
  characteristics
\item
\item
  Dominance of weak interactions in concordance with other studies.
\end{enumerate}

Within a system of a given complexity, commonly only a few strong
species interactions are present with most interactions being weak
(e.g.~Bascompte et al., 2005; Berlow et al., 2004; Paine, 1992; Wootton
\& Emmerson, 2005), though still being important for system stability
(e.g.~Kadoya et al., 2018; McCann et al., 1998; O'Gorman \& Emmerson,
2009). Moreover, there is some evidence indicating that generalists
(i.e.~species with many possible interacting partners) mostly have weak
interactions, while specialist species (i.e.~fewer possible interacting
partners) show stronger interactions (Wootton \& Stouffer, 2016).

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
\item
\end{enumerate}

Our results contradict studies suggesting that mid-trophic level species
are key in marine polar ecosystems. \citep{Murphy2016}, studies by
Australians

>>>>

Our results regarding the distribution of the interaction strengths are
consistent with prior theoretical and empirical studies
\citep[e.g.][]{McCann1998, Neutel2002, Emmerson2004, Wootton2005}.
Within a system of a given complexity, commonly only a few strong
species interactions are present with most interactions being weak.
Historically, after \citet{May1972} installed the debate
complexity-stability, the results about the distribution of the
interaction strength has been interpreted as an explanation for the
persistence of complex food webs in nature
\citep{Bascompte2005, Allesina2015, Nilsson2016}. Here we showed that
this pattern is also prevalent in one of the most complex empirical
marine food webs that was studied so far, with 490 species and 16041
predator-prey interactions.

Our approach to study species properties follows \citet{Kortsch2021}
recommendation that suggests to employ a range of descriptors from both
unweighted and weighted food web in order to characterise the dynamic
and multifaceted nature of structural and functional ecosystem changes.
We went a step forward and analysed the relationship between weighted
and unweighted descriptors. In this regard, our results show a positive
relationship between interaction strength (weighted) and trophic level
(unweighted), and between interaction strength and degree (total number
of interactions). The former relationship might contradict those studies
that suggest that mid-trophic level species are involved in the major
pathways of energy flow in high-latitude marine ecosystems
\citep{Pinkerton2014, Murphy2016, McCormack2020, Riccialdelli2020}. This
could be explained by the lack of species biomass data in our
interaction strength estimations; the methodology we applied here
\citep{Pawar2012} allows to include empirical data for the density of
the resource, though not for all food web species. Unfortunately, the
lack of individual data for the entire Weddell Sea food web hampers any
alternative approach. On the other hand, the positive relationship
between interaction strength and degree reinforces the importance of
species with many interactions: species with high degree (hubs) have a
large impact on overall food web structure and functioning
\citep{Dunne2002a, Kortsch2015}. In the Weddell Sea, species with high
degree also tend to have high mean interaction strengths. This
information on the quantity and quality of interactions and its
relationship enables a robust assessment of the species' role in the
stability of the food web \citep{Cirtwill2018a}.

Our results show that only few species are key in terms of stability in
the Weddell Sea food web. This is in concordance with studies of complex
empirical food webs in marine ecosystems in the Arctic and other
locations in Antarctica \citep{Kortsch2015, Marina2018, Rodriguez2022}.
In the present work we identify that such key species present particular
weighted and unweighted properties: 1) high mean interaction strength;
2) mid to high trophic level; 3) relatively high number of interactions;
and 4) mid to low trophic similarity. The first attempt to gain insight
into the role of species in the Weddell Sea applying a complex approach
was that of \citet{Jacob2011}. These authors used information on body
size, predatory and feeding type, and environment (habitat) to study
species' roles. After performing sequential extinction simulations,
\citet{Jacob2011} found that larger bodied species can be lost without
causing a direct collapse of the network topology. A major caveat of
this finding, already recognized by the authors, is that population
dynamics were ignored and therefore no top-down extinctions, or other
indirect effects, could occur. In our study we did consider such effects
by including information on the species interaction strength, which is
of paramount importance when analysing the response of perturbations in
ecological communities \citep{McCann1998, Montoya2009, Novak2011}. Thus,
our study suggests that species with high mean interaction strength and
high trophic level need to be considered with particular attention when
trying to predict the effects of perturbations on the Weddell Sea
ecosystem. Furthermore, this arises as a greater issue since we have
identified that these species present a mid to low trophic similarity,
meaning that few other species of the food web can occupy the same
trophic role. In this sense, \citet{Murphy2016}'s review emphasized that
polar pelagic ecosystems are particularly sensitive to change due to a
low functional redundancy at key trophic levels. It's important to note
that we provide a broader analysis in this regard because we have
included species from all habitats (pelagic, benthic, land-based).
